### 2021-11-17 plos_latex_template.tex
Changes to section structure

\section{Embrace all dimensions of diversity}
\label{rule_diversity}

@@

\section{Create a safe and welcoming environment}
\label{rule_inclusion}

@@

\section{Have an inclusive and diverse organizing team}
\label{rule_organizing_team}

@@

A genuinely inclusive conference can only be coordinated  by an inclusive and diverse organizing team, in charge of the conference decision-making.
Organizers should build a team with people from different regions, genders, races and ethnicities, socioeconomic statuses, and other aspects of diversity.
Having diverse people in decision-making positions will affect positively all the other aspects of your conference because all the processes will benefit from their input, expertise, and distinct perspectives \cite{hongGroupsDiverseProblem2004}. 
In addition, a diverse team plays an important role at creating a welcoming space because representation--seeing people with similar life experiences occupy public spaces, positions of power, and breaking negative stereotypes--is one of the best ways to create a sense of belonging for everyone participating in the conference (see \textbf{Rule \ref{rule_inclusion}}). If you already started assembling an organizing team, check for gaps in its composition. 
The regional and local communities, groups, or associations in your field are good sources to tap into.

@@

\section{Consciously counteract bias in the conference program}
\label{rule_unbias}
%la: LA: repeating parts of my comment in Figure 1:maybe "mitigate bias" - unbiasedness is a reasonable goal once systemic biases are gone, in the meantime, introducing biases favoring marginalized folks is required to reach equity. Thus, while unbiasedness is the ultimate goal, the current rules should point to mitigate biases and that may require to bias in favor of marginalized groups. Equity and Justice call for biases in favor of marginalized folks. Unbiasing aims for equality that, in the current status quo, will keep favoring those who are already in the mainstream and will keep increasing marginalization for those currently in the margins. This is a great reference regarding the framework "mainstream/margins" https://www.trainingforchange.org/wp-content/uploads/2017/11/Mainstream-Margin-in-Groups.pdf See the cartoon about equality, equity, and justice here: https://onlinepublichealth.gwu.edu/resources/equity-vs-equality/ and here: https://www.reddit.com/r/coolguides/comments/o3ivsg/equality_equity_and_justice_explained_better/
%idea of paragraph 1: our lists are biased

@@

\section{Have a strong online component} 
\label{rule_online}

@@

\section{Make the conference accessible to people with disabilities}
\label{rule_accessibility}

Conferences are among the least accessible spaces that people with disabilities may encounter in professional contexts \cite{priceAccessImaginedConstruction2009}. Even when conferences implement other inclusive practices, the participation of people with disabilities is often overlooked \cite{marks2021meeting}. This is one key aspect where participation in the organizing team allows disabled people to take part in the decisions from the beginning (see \textbf{Rule \ref{rule_organizing_team}}), as thinking about disability or simulating it are not substitutes for real-life experience \cite{costanzachockDesign2020}. Planning for accessibility requires time and early decision-making \cite{irishIncreasingParticipationUsing2020}. As other aspects of inclusion, dealing with accessibility at the last minute is the recipe for a disastrous conference. If you do not consider accessibility from the conference inception, it is highly likely you will be better off without trying to patch accessibility at the last minute. Key decisions in this respect are hard to correct. 
For in-person conferences, the venue should comply with common accessibility standards, such as being adequate for people who use wheelchairs, have signs in Braille, and a sound system compatible with hearing devices and live interpretation, just to name a few important features. In addition to this, the organizers should take care proactively of invisible disabilities (e.g., dyslexia, anxiety, ADHD, autism and its spectrum). This could be done in many ways, for example, by providing quiet spaces for privacy and noise-free conversations, provide chairs in open spaces, and menus that menus that account for the dietary requirements participants need to respect.%ast: check the flow here

@@

Accessibility practices should also include social events and networking, and include activities that do not restrict participation based on body type or ability. Importantly, all these accessibility practices are inclusive not only for people with disabilities but to everyone.
For instance, captions are helpful for non-native speakers, having the material available for download helps attendees with low bandwidth connection, etc. Accessibility efforts should be explicitly displayed in the official communication channels  (see \textbf{Rule \ref{rule_communication}}).

\section{Don't let language restrict high-quality participation}
\label{rule_language}

@@

% Rocío: main idea: make conferences more linguistically inclusive
Providing a welcoming and diverse environment by encouraging the full participation of non-native English speakers is critical (Rule \ref{rule_inclusion}). This may be done at different levels. First, by having spaces in languages other than English, both regarding the technical content of the conference and the social and networking aspects. These exclusive spaces have the benefit of not requiring technical adjustments, but have the problem of not communicating outside their specific public.   
Second, for presentations spoken in English, identify some other key languages for the conference and provide at least English-to-English captions (see Rule \ref{rule_accessibility}), but also translated captions or live interpretation into these key languages. Allow for multilingual Q \& A sessions with an interpreter. Third, include workshops, presentations, and events in languages other than English in the schedule, and provide captions or live interpretation to English.

@@

\section{Express the welcoming spirit in your communication strategy}
% media
\label{rule_communication}

%Rocío: Main idea: Include people in your communication 
Your communication strategy should reflect the spirit of the conference and be inclusive by design. 
As you design the promotion plan, actively reach out and promote the conference to people who have been systematically excluded. 
Include all the work to make the conference a safe and inclusive environment in your communication strategy, 
to emphasize that everybody is welcome in this event. This could be done through publishing a diversity statement (e.g., https://user2021.r-project.org/about/diversitystatement/), the Code of Conduct, the accessibility guidelines, and advertising financial support (see Rule \ref{rule_financial}). Try to come up with creative ways to show that everyone is seen, respected, and welcome. \la{For example, useR! 2021 created "Margot, the marmot" a mascot for the conference. Margot is a cartoon with different designs reflecting folks that the conference wanted to reach. In particular, Margot wears different scarfs. The scarfs include colors of flags from different communities in R such as MiR´s, Rainbow R, ..., etc. (Figure 2)."}.
Also, try to advertise the conference in multiple languages (\textbf{Rule \ref{rule_language}}) and multiple platforms (e.g. Twitter, Facebook, LinkedIn, conference website, mailing lists)--
part of your community may use a different platform than the ones adopted by people who regularly attend the conference. 
% rj: allude to the diverse organizing team - to tap into groups, contexts and make choices on platforms and ways to communicate
With proactive inclusive communication strategies, potential attendees will be assured that this is their space and community too.

@@

\section{Allocate adequate financial resources to support your conference goals}
\label{rule_financial}

@@

\section{Make the conference part of a long-term process for inclusion}
\label{rule_process}

>>>>

\section*{The pillars}

To make a more inclusive conference, it is necessary to identify which groups of people are being excluded, the barriers they may face to be part of the conference, and strategies to reduce them. 
The first rule is about embracing all dimensions of diversity and prioritizing the
 participation of groups that are often marginalized.
However, increasing diversity is not enough. Rule 2 is about creating a safe and welcoming environment to make everyone feel part of the conference and the community. 
Inclusion should start from the organizing team. Rule 3 refers to the importance of building an inclusive and diverse organizing team, and provides advice on work dynamics.

\subsection*{Rule 1: Embrace all dimensions of diversity}
\label{rule_diversity}

@@

\subsection*{Rule 2: Create a safe and welcoming environment}
\label{rule_inclusion}

@@

\subsection*{Rule 3: Have an inclusive and diverse organizing team}
\label{rule_organizing_team}

@@

A genuinely inclusive conference can only be coordinated  by an inclusive and diverse organizing team, in charge of the conference decision-making.
Organizers should build a team with people from different regions, genders, races and ethnicities, socioeconomic statuses, and other aspects of diversity.
Having diverse people in decision-making positions will affect positively all the other aspects of your conference because all the processes will benefit from their input, expertise, and distinct perspectives \cite{hongGroupsDiverseProblem2004}. 
In addition, a diverse team plays an important role at creating a welcoming space because representation--seeing people with similar life experiences occupy public spaces, positions of power, and breaking negative stereotypes--is one of the best ways to create a sense of belonging for everyone participating in the conference (see Rule 2 about a safe and welcoming environment). If you already started assembling an organizing team, check for gaps in its composition. 
The regional and local communities, groups, or associations in your field are good sources to tap into.

@@

\section*{Design Rules}

These rules concern diverse elements in the design of the conference. They are all important pieces for inclusion.
In rule 4, we introduce ways to counteract bias in the conference program, in keynotes, program committee, abstract selection, and other spaces. 
Rule 5 describes the advantages of online conferences reducing barriers for participation, and provides advice for organizing online and hybrid conferences.
Rule 6 focuses on practices to make all conference platforms and spaces accessible to people with disabilities. 
In rule 7, we advocate for more language-inclusive conferences, and provide suggestions to encourage and ensure full participation of non-native English speakers. 
Rule 8 provides tips for an inclusive media strategy. 
In rule 9, we address budgeting for inclusive practices, and helping participants with affordable registration costs, scholarships, and other forms of financial support.

\subsection*{Rule 4: Consciously counteract bias in the conference program}
\label{rule_unbias}
%la: LA: repeating parts of my comment in Figure 1:maybe "mitigate bias" - unbiasedness is a reasonable goal once systemic biases are gone, in the meantime, introducing biases favoring marginalized folks is required to reach equity. Thus, while unbiasedness is the ultimate goal, the current rules should point to mitigate biases and that may require to bias in favor of marginalized groups. Equity and Justice call for biases in favor of marginalized folks. Unbiasing aims for equality that, in the current status quo, will keep favoring those who are already in the mainstream and will keep increasing marginalization for those currently in the margins. This is a great reference regarding the framework "mainstream/margins" https://www.trainingforchange.org/wp-content/uploads/2017/11/Mainstream-Margin-in-Groups.pdf See the cartoon about equality, equity, and justice here: https://onlinepublichealth.gwu.edu/resources/equity-vs-equality/ and here: https://www.reddit.com/r/coolguides/comments/o3ivsg/equality_equity_and_justice_explained_better/
%idea of paragraph 1: our lists are biased

@@

\subsection*{Rule 5: Have a strong online component} 
\label{rule_online}

@@

\subsection*{Rule 6: Make the conference accessible to people with disabilities}
\label{rule_accessibility}

Conferences are among the least accessible spaces that people with disabilities may encounter in professional contexts \cite{priceAccessImaginedConstruction2009}. Even when conferences implement other inclusive practices, the participation of people with disabilities is often overlooked \cite{marks2021meeting}. This is one key aspect where participation in the organizing team allows disabled people to take part in the decisions from the beginning (see rule 3 about the organizing team), as thinking about disability or simulating it are not substitutes for real-life experience \cite{costanzachockDesign2020}. Planning for accessibility requires time and early decision-making \cite{irishIncreasingParticipationUsing2020}. As other aspects of inclusion, dealing with accessibility at the last minute is the recipe for a disastrous conference. If you do not consider accessibility from the conference inception, it is highly likely you will be better off without trying to patch accessibility at the last minute. Key decisions in this respect are hard to correct. 
For in-person conferences, the venue should comply with common accessibility standards, such as being adequate for people who use wheelchairs, have signs in Braille, and a sound system compatible with hearing devices and live interpretation, just to name a few important features. In addition to this, the organizers should take care proactively of invisible disabilities (e.g., dyslexia, anxiety, ADHD, autism and its spectrum). This could be done in many ways, for example, by providing quiet spaces for privacy and noise-free conversations, provide chairs in open spaces, and menus that menus that account for the dietary requirements participants need to respect.%ast: check the flow here

@@

Accessibility practices should also include social events and networking, and include activities that do not restrict participation based on body type or ability. Importantly, all these accessibility practices are inclusive not only for people with disabilities but to everyone.
For instance, captions are helpful for non-native speakers, having the material available for download helps attendees with low bandwidth connection, etc. Accessibility efforts should be explicitly displayed in the official communication channels (see rule 8 about media strategy).

\subsection*{Rule 7: Don't let language restrict high-quality participation}
\label{rule_language}

@@

% Rocío: main idea: make conferences more linguistically inclusive
Providing a welcoming and diverse environment by encouraging the full participation of non-native English speakers is critical (rule 2 about a welcoming environment for all). This may be done at different levels. First, by having spaces in languages other than English, both regarding the technical content of the conference and the social and networking aspects. These exclusive spaces have the benefit of not requiring technical adjustments, but have the problem of not communicating outside their specific public.   
Second, for presentations spoken in English, identify some other key languages for the conference and provide at least English-to-English captions (see rule 6 about accessibility for disabilities), but also translated captions or live interpretation into these key languages. Allow for multilingual Q \& A sessions with an interpreter. Third, include workshops, presentations, and events in languages other than English in the schedule, and provide captions or live interpretation to English.

@@

\subsection*{Rule 8: Express the welcoming spirit in your communication strategy}
% media
\label{rule_communication}

%Rocío: Main idea: Include people in your communication 
Your communication strategy should reflect the spirit of the conference and be inclusive by design. 
As you design the promotion plan, actively reach out and promote the conference to people who have been systematically excluded. 
Include all the work to make the conference a safe and inclusive environment in your communication strategy, 
to emphasize that everybody is welcome in this event. This could be done through publishing a diversity statement (e.g., https://user2021.r-project.org/about/diversitystatement/), the Code of Conduct, the accessibility guidelines, and advertising financial support (see rule 9 about financial resources). Try to come up with creative ways to show that everyone is seen, respected, and welcome. \la{For example, useR! 2021 created "Margot, the marmot" a mascot for the conference. Margot is a cartoon with different designs reflecting folks that the conference wanted to reach. In particular, Margot wears different scarfs. The scarfs include colors of flags from different communities in R such as MiR´s, Rainbow R, ..., etc. (Figure 2)."}.
Also, try to advertise the conference in multiple languages (see rule 7 about language) and multiple platforms (e.g. Twitter, Facebook, LinkedIn, conference website, mailing lists)--
part of your community may use a different platform than the ones adopted by people who regularly attend the conference. 
% rj: allude to the diverse organizing team - to tap into groups, contexts and make choices on platforms and ways to communicate
With proactive inclusive communication strategies, potential attendees will be assured that this is their space and community too.

@@

\subsection*{Rule 9: Allocate adequate financial resources to support your conference goals}
\label{rule_financial}

@@

\section*{The Process}

One conference can make a difference in the path towards inclusion. The organizers of a conference edition and selection committees should do their best to pass their knowledge on best practices and problems encounters to the next organizers.

\subsection*{Rule 10: Make the conference part of a long-term process for inclusion}
\label{rule_process}
